\newpage
\chapter*{Liste de publications et/ou brevets et participation aux conférences}

\addcontentsline{toc}{chapter}{List of publications and conference participation}
\subsection*{Liste des publications et/ou brevet réalisés dans le cadre du projet de thèse :}


\begin{enumerate}
\item \textbf{Ladret, H. J.}, Casanova, C., \& Perrinet, L. U., (2023). \textit{Kernel Heterogeneity Improves Sparseness of Natural Images Representations}. preprint available as an arXiv preprint arXiv:2312.14685, under review (as of 01/01/2024) in Neuromorphic Computing and Engineering.


\item Cortes, N., \textbf{Ladret, H. J.}, Abbas Farishta, R., \& Casanova, C. (2023). \textit{The pulvinar as a hub of visual processing and cortical integration}. published in Trends in Neurosciences, dec 2023. 

DOI:https://doi.org/10.1016/j.tins.2023.11.008


\item \textbf{Ladret, H. J.}, Cortes, N., Ikan, L., Chavane, F., Casanova, C., \& Perrinet, L. U. (2023). \textit{Cortical recurrence supports resilience to sensory variance in the primary visual cortex}. published in Communications Biology, 6(1), 667. 

DOI:https://doi.org/10.1038/s42003-023-05042-3


\item Cortes, N., Abbas Farishta, R., \textbf{Ladret, H. J.}, \& Casanova, C. (2021). \textit{Corticothalamic Projections Gate Alpha Rhythms in the Pulvinar}. published in Frontiers in Cellular Neuroscience, 15, 787170. 

DOI:https://doi.org/10.3389/fncel.2021.787170
\end{enumerate}


\subsection*{Participation aux conférences et écoles d'été au cours de la période de thèse :}

\subsubsection*{Articles de conférences :}
\begin{enumerate}
\item \textbf{Ladret, H. J.}, Perrinet, L. U., \& Casanova, C. (2023). \textit{Convolutional Sparse Coding is improved by heterogeneous uncertainty modeling}. Proceedings of ICLR SNN, Kigali, Rwanda.
\item \textbf{Ladret, H. J.}, Cortes, N., Ikan, I., Chavane, F., Casanova, C., \& Perrinet, L. U. (2023). \textit{Resilience to sensory uncertainty in the primary visual cortex}. Proceedings of Cosyne, Montreal, Canada. 
\end{enumerate}

\subsubsection*{Présentations orales :}
\begin{enumerate}
\item \textbf{Ladret, H. J.} (2022). \textit{Making sense of the visual mess}. Presented at Aix-Marseille's doctoral schools summit, Marseille, France.
\item \textbf{Ladret, H. J.} (2022). \textit{Dealing with sensory variance in the primary visual cortex}. Invited talk at Netherlands Institute of Neuroscience, Amsterdam, Netherlands.
\item \textbf{Ladret, H. J.}, Perrinet, L. U. (2022). \textit{Statistics of the sparse representations of natural images}. Presented at SIAM Conference, Virtual.
\item \textbf{Ladret, H. J.} (2021). \textit{Modulation of orientation selectivity by orientation precision in V1}. Presented at GDR Vision, Lille, France. 
\item \textbf{Ladret, H. J.}, Perrinet, L. U. (2021). \textit{Dynamics of the processing of orientation precision in the primary visual cortex}. Invited talk at DynamicsDays, Sophia-Antipolis.
\end{enumerate}

\subsubsection*{Présentations sous forme de poster :}
\begin{enumerate}
\item \textbf{Ladret, H. J.}, Cortes, N., Ikan, I., Chavane, F., Casanova, C. \& Perrinet, L. U. (2023). \textit{Computing sensory variance through intracortical recurrence}. Poster session presented at VSS, Tampa, USA.
\item \textbf{Ladret, H. J.}, Cortes, N., Ikan, I., Chavane, F., Casanova, C. \& Perrinet, L. U. (2023). \textit{Resilience to sensory uncertainty in the primary visual cortex}. Poster session presented at Cosyne, Montreal, Canada. 
\item \textbf{Ladret, H. J.} \& Perrinet, L. U. (2022). \textit{Uncertainty in, uncertainty out: epistemic variance improves encoding of natural images}. Poster session presented at GDR Vision, Toulouse, France.
\item Ikan, I., Cortes, N., \textbf{Ladret, H. J.}, Laplante, I. \& Casanova, C. (2022). \textit{Dynamics of response’s accuracy in the visual cortical area 21a}. Poster session presented at SfN, San Diego, USA. 
\item \textbf{Ladret, H. J.}, Cortes, N., Ikan, I., Chavane, F., Casanova, C. \& Perrinet, L. U. (2022). \textit{Input variance and V1}. Poster session presented at INT X Anniversary, Marseille, France.
\item Bourquin, N., Perrot, A., Porée, L., Belgharbi, S., Cortes, N., Miquel, C., Bélanger, E., \textbf{Ladret, H. J.}, Ikan, I., Thorin, C., Lesage, F. \& Provost, J. (2022). \textit{Dynamic Ultrasound Localisation Microscopy Achieves Quantitative pulsatility Measurements in the Whole Brain Using Kalman Filtering}. Poster session presented at IEEE US, Virtual.
\item \textbf{Ladret, H. J.}, Cortes, N., Ikan, I., Chavane, F., Casanova, C. \& Perrinet, L. U. (2022). \textit{Recurrent cortical connectivity in the primary visual cortex supports robust encoding of natural sensory inputs}. Poster session presented at FENS, Paris, France. 
\item \textbf{Ladret, H. J.}, \& Perrinet, L. U. (2022). \textit{A resilient neural code in V1 to process natural images}. Poster session presented at AREADNE, Santorini, Greece. 
\item \textbf{Ladret, H. J.}, Cortes, N., Ikan, I., Chavane, F., Casanova, C. \& Perrinet, L. U. (2021). \textit{Modulation of orientation selectivity by orientation precision}. Poster session presented at SfN, Virtual. 
\item \textbf{Ladret, H. J.}, \& Perrinet, L. U. (2021). \textit{Decoding orientation distributions from noisy observations in the primary visual cortex}. Poster session presented at Champalimaud's Dialogues on Neural and Machine Intelligence, Lisbon, Portugal.
\item \textbf{Ladret, H. J.}, Cortes, N., Ikan, I., Chavane, F., Casanova, C. \& Perrinet, L. U. (2021). \textit{Processing of orientation precision in the primary visual cortex}. Poster session presented at NeuroFrance Annual Meeting, Strasbourg, France.
\item \textbf{Ladret, H. J.}, Cortes, N., Perrinet, L. U. \& Casanova, C. (2019). \textit{Orientation selectivity to synthetic natural patterns in a cortical-like model of the cat primary visual cortex}. Poster session presented at SfN, Chicago, USA.
\item \textbf{Ladret, H. J.} \& Ibos, G. (2019). \textit{Comparative decision-making in the Posterior Parietal Cortex: proof of concept}. Poster session presented at SfN, Chicago, USA.
\item \textbf{Ladret, H. J.}, Cortes, N., Casanova, C., \& Perrinet, L. U. (2019). \textit{Learning dynamics in a neural network model of the primary visual cortex}. Poster session presented at Vision Health Research Network, Quebec, Canada. 
\item \textbf{Ladret, H. J.} \& Perrinet, L. U. (2018). \textit{Selectivity to oriented patterns of different precisions}. Poster session presented at GDR Vision, Paris, France. 
\end{enumerate}

\subsubsection*{Écoles d'été :}
\begin{enumerate}
\item (2023) \textit{Telluride Neuromorphic Cognition Engineering Workshop}. Project leader on a brain organoid / open-source software interface project. Telluride, Colorado, USA.
\item (2022) \textit{European Summer School of Vision}. Schloss Rauischholzhausen, Hessia, Germany.
\end{enumerate}